A Formal Interface for Concurrent Search Structure Templates

### 2025-04-28 concurrency/template paper/paper.tex
sketched implementation layout

\section{Coding Search Structure Templates}
\begin{figure}
  \begin{subfigure}[t]{.4\textwidth}
  \begin{lstlisting}[language=C, style=myStyle]
  struct node{ int key; int value;
    node* next;
    lock l; }
  \end{lstlisting}
    \caption{Linked-list node with lock}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{.4\textwidth}
  \begin{lstlisting}[language=C, style=myStyle]
  struct node{ int key; int value;
    node* next;
    lock l; int min; int max; }
  \end{lstlisting}
    \caption{Linked-list node with lock and range}
  \end{subfigure}

@@

\section{Evaluation}

@@

>>>>

\section{Coding Search Structure Templates}
\label{implementation}
\begin{figure}
  \begin{subfigure}[t]{.4\textwidth}
  \begin{lstlisting}[language=C, style=myStyle]
  struct node{ int key; int value;
    node* next;
    lock l; }
  \end{lstlisting}
    \caption{Linked-list node with lock}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{.4\textwidth}
  \begin{lstlisting}[language=C, style=myStyle]
  struct node{ int key; int value;
    node* next;
    lock l; int min; int max; }
  \end{lstlisting}
    \caption{Linked-list node with lock and range}
  \end{subfigure}

@@

@@

\section{Implementation and Evaluation}
The code for our concurrent data structures is written in C, and our specifications and proofs are written in the Verified Software Toolchain (VST)~\cite{plcc}. Using C forces us to confront the implementation challenges described in section~\ref{implementation}, which we might accidentally circumvent in a core calculus. VST allows us to prove separation logic specifications of C programs in Rocq, and newer versions of VST~\cite{iris-compcert} are built on Iris, allowing us to directly reuse the implementation of flow interfaces from \citet{templates}. Our C code and our Rocq proofs follow a common structure that ensures modularity. %diagram:                       %should we put some of this structural stuff earlier?
%                   data_structure.h
%             /              |                 \
%         bst.c          list.c             template.h
%                                       /            |             \
%                                give-up.c   coupling.c   template.c

The files \texttt{data\_structure.h} and \texttt{template.h} define the interfaces for data structures and templates respectively, and the corresponding VST files define the typeclasses for these interfaces, including the specification of each interface function. \texttt{bst.c} and \texttt{list.c} implement the data structure interface, and \texttt{give-up.c} and \texttt{coupling.c} implement the template interface. Finally, \texttt{template.c} implements the top-level operations, relying on the interfaces for data structures and templates but not on any details of their implementation (and the corresponding VST file verifies the operations with the interface typeclasses as parameters). Importantly, there is no file in either C or Rocq that is specific to a combination of data structure and template, such as BST+give-up or list+coupling: we can freely combine any data structure and concurrency template and obtain a working concurrent data structure with no further effort, simply by compiling \texttt{template.c} along with the chosen data structure and template file.

(LoC/LoP numbers for each part)
(comparison of proof size to acsys proofs and/or CPP paper?)
